How phylogenetic relatedness and floral traits are involved in heterospecifc pollen effect in an artificial co-flowering community

### 2018-11-02 Manuscript_draft/Manuscript_draft.Rmd
other changes...
adding some refs

---
title: 'How Phylogeny and traits relate to heterospecific pollen effect'

@@

Write your abstract here.

@@

*Keywords*: rmarkdown, reproducible science

@@

@@

Foreign pollen effect was studied through two different treatments, one with 50% conspecific pollen and 50% heterospecific pollen and a second one with 100% foreign pollen (N=10). Moreover, hand cross pollination, hand self pollination, apomixis (bagged emasculated flowers) and natural selfing were tested (N=10). Flowers were emasculated the day prior anthesis and hand pollinated next day with a toothpick. Had-pollination was realized with 3-4 gentle touches on the surface of the stigma. The mixes of pollen were performed based on the pollen counts (Neubaeur chamber) on an eppendorf.

Explain the phylogeny...

Explain the measurement of the traits...

@@

Trees in forest A grew taller than those in forest B (mean height: `r mean(25, 31, 28)` versus `r mean(13, 19, 16)` m). And many more cool results that get updated dynamically.

>>>>

---
title: 'How phylogenetic relatedness and floral traits are involved in heterospecifc pollen effect in an artificial co-flowering community'

@@

Possible journals to publish: New phytologist, journal of ecology, oikos...

@@

*Keywords*: heterospecific pollen, plant reproduction, fitness, competition

@@

**Paragraph 1**

In natural systems plant species normally coexist and share their floral visitors with other species (Bascompte et al., 2003). This pollinator sharing from the plant perpective can be negative due to competition (refs) or positive due to facilitation (refs). Moreover, once the pollinator has landed on the stigma some other issues for the species fitness may arise, the arrival of foreign pollen and conspecific pollen loss.

**Paragraph 2**

The effect of heterospecific pollen has been widely studied (refs). Native species are supposed to have greater negative effects than native ones @arceo2016. However, the quantity of pollen that integrates in the network can be quite low @bartomeus2008. Moreover, closely related species are supposed to reduce fitness in greter effect but the evidence is scarce and based on independent studies with different methodlogies (refs) or studies that just check that with very highly close related species with the aim of study hybridization (refs).

**Paragraph 3**

Relatedness. What we know?

**Paragraph 4**

The great difficulty of working with pollen in a coflowering community make the understanding of heterospecifc pollen effect a real challenge. For this reason we have created an artificial co-flowering community in a glasshouse to test the effect with all the possible combinations among them. Where we test the folowing hypothesis: 1) Does heterospecific pollen reduce seed set, if so, 2) Does heterospecific pollen effect depend on the relatedness of the species, 3) Does heterospecific pollen effect depend on any floral trait?

@@

Foreign pollen effect was studied through two different treatments, one with 50% conspecific pollen and 50% heterospecific pollen and a second one with 100% foreign pollen (N=10). The proxy of effect were seed set and "pollen tubes". Moreover, hand cross pollination, hand self pollination, apomixis (bagged emasculated flowers) and natural selfing were tested (N=10). Flowers were emasculated the day prior anthesis and hand pollinated next day with a toothpick. Had-pollination was realized with 3-4 gentle touches on the surface of the stigma. The mixes of pollen were performed on an eppendorf based on the pollen counts maded with Neubaeur chamber (each anther was counted 4 times for 20 different anthers per species).

**Evolutive distance**

Two types of evolutive distances were calculated with MEGA7 thow kinds of markers: 1) Internal transcribed spacer (ITS) and 2) ribulose-bisphosphate carboxylase (RBCL)

**Traits**

Several traits of the ten species were measured. Pollen per anther was counted, number of ovules, stigma width and length and stigmatic area, style width and length, ovary width and length. Moreover stigma type was tested. Self-incompatibility was

@@

### 2018-11-02 Manuscript_draft/Manuscript_draft.tex
adding some refs

\documentclass[11pt,a4paper]{article}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
    \usepackage{xltxtra,xunicode}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\usepackage[margin=1in]{geometry}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={How Phylogeny and traits relate to heterospecific pollen effect},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\setcounter{secnumdepth}{0}

@@

\title{How Phylogeny and traits relate to heterospecific pollen effect}
\author{
Jose B. Lanuza
true
}
\date{}
% Allowing for landscape pages
\usepackage{lscape}
\newcommand{\blandscape}{\begin{landscape}}
\newcommand{\elandscape}{\end{landscape}}

@@

\begin{document}
%%Edited by @jhollist 09/15/2014
%%Adds title from YAML
\begin{singlespace}
\begin{center}
\huge How Phylogeny and traits relate to heterospecific pollen effect
\end{center}
%%Adds Author, correspond email asterisk, and affilnum from YAML
\begin{center}
\large
Jose B. Lanuza \textsuperscript{1,2} 
true
\end{center}
%%Adds affiliations from YAML
\begin{justify}
\footnotesize \emph{ 
\\*
\textsuperscript{1}US Environmental Protection Agency, Office of Research and Development,
National Health and Environmental Effects Research Laboratory, Atlantic
Ecology Division, 27 Tarzwell Drive Narragansett, RI, 02882, USA\\*
\\*
\textsuperscript{2}Big Name University, Department of R, City, BN, 01020, USA\\*
\\*
\textsuperscript{3}Estacion Biologica de Donana (EBD-CSIC), E-41092 Sevilla, Spain\\*
}
%%Adds corresponding author email(s) from YAML
\newcounter{num}
\setcounter{num}{1}
\\[0.1cm]
\footnotesize \emph{ 
\ifnum\value{num}=1%
\textsuperscript{*} corresponding author:
\fi
\href{mailto:f.rodriguez.sanc@gmail.com}{\nolinkurl{f.rodriguez.sanc@gmail.com}}
\stepcounter{num}
}
\end{justify}
%%Adds date from YAML
\normalsize

@@

Write your abstract here.

@@

\emph{Keywords}: rmarkdown, reproducible science

@@

@@

Foreign pollen effect was studied through two different treatments, one
with 50\% conspecific pollen and 50\% heterospecific pollen and a second
one with 100\% foreign pollen (N=10). Moreover, hand cross pollination,
hand self pollination, apomixis (bagged emasculated flowers) and natural
selfing were tested (N=10). Flowers were emasculated the day prior
anthesis and hand pollinated next day with a toothpick. Had-pollination
was realized with 3-4 gentle touches on the surface of the stigma. The
mixes of pollen were performed based on the pollen counts (Neubaeur
chamber) on an eppendorf.

@@

Trees in forest A grew taller than those in forest B (mean height: 25
versus 13 m). And many more cool results that get updated dynamically.

@@

>>>>

\documentclass[11pt,a4paper]{article}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
    \usepackage{xltxtra,xunicode}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\usepackage[margin=1in]{geometry}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={How phylogenetic relatedness and floral traits are involved in heterospecifc pollen effect in an artificial co-flowering community},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\setcounter{secnumdepth}{0}

@@

\title{How phylogenetic relatedness and floral traits are involved in
heterospecifc pollen effect in an artificial co-flowering community}
\author{
Jose B. Lanuza
true
}
\date{}
% Allowing for landscape pages
\usepackage{lscape}
\newcommand{\blandscape}{\begin{landscape}}
\newcommand{\elandscape}{\end{landscape}}

@@

\begin{document}
%%Edited by @jhollist 09/15/2014
%%Adds title from YAML
\begin{singlespace}
\begin{center}
\huge How phylogenetic relatedness and floral traits are involved in
heterospecifc pollen effect in an artificial co-flowering community
\end{center}
%%Adds Author, correspond email asterisk, and affilnum from YAML
\begin{center}
\large
Jose B. Lanuza \textsuperscript{1,2} 
true
\end{center}
%%Adds affiliations from YAML
\begin{justify}
\footnotesize \emph{ 
\\*
\textsuperscript{1}US Environmental Protection Agency, Office of Research and Development,
National Health and Environmental Effects Research Laboratory, Atlantic
Ecology Division, 27 Tarzwell Drive Narragansett, RI, 02882, USA\\*
\\*
\textsuperscript{2}Big Name University, Department of R, City, BN, 01020, USA\\*
\\*
\textsuperscript{3}Estacion Biologica de Donana (EBD-CSIC), E-41092 Sevilla, Spain\\*
}
%%Adds corresponding author email(s) from YAML
\newcounter{num}
\setcounter{num}{1}
\\[0.1cm]
\footnotesize \emph{ 
\ifnum\value{num}=1%
\textsuperscript{*} corresponding author:
\fi
\href{mailto:barragansljose@gmail.com}{\nolinkurl{barragansljose@gmail.com}}
\stepcounter{num}
}
\end{justify}
%%Adds date from YAML
\normalsize

@@

Possible journals to publish: New phytologist, journal of ecology,
oikos\ldots{}

@@

\emph{Keywords}: heterospecific pollen, plant reproduction, fitness,
competition

@@

\textbf{Paragraph 1}

In natural systems plant species normally coexist and share their floral
visitors with other species (Bascompte et al., 2003). This pollinator
sharing from the plant perpective can be negative due to competition
(refs) or positive due to facilitation (refs). Moreover, once the
pollinator has landed on the stigma some other issues for the species
fitness may arise, the arrival of foreign pollen and conspecific pollen
loss.

\textbf{Paragraph 2}

The effect of heterospecific pollen has been widely studied (refs).
Native species are supposed to have greater negative effects than native
ones Arceo-Gómez and Ashman (2016)

species-specific. What kind of pollen interaction? Size just mechanical?

\textbf{Paragraph 3}

Relatedness. What we know?

\textbf{Paragraph 4}

The great difficulty of working with pollen in a coflowering community
make the understanding of heterospecifc pollen effect a real challenge.
For this reason we have created an artificial co-flowering community in
a glasshouse to test the effect with all the possible combinations among
them. Where we test the folowing hypothesis: 1) Does heterospecific
pollen reduce seed set, if so, 2) Does heterospecific pollen effect
depend on the relatedness of the species, 3) Does heterospecific pollen
effect depend on any floral trait?

@@

Foreign pollen effect was studied through two different treatments, one
with 50\% conspecific pollen and 50\% heterospecific pollen and a second
one with 100\% foreign pollen (N=10). The proxy of effect were seed set
and ``pollen tubes''. Moreover, hand cross pollination, hand self
pollination, apomixis (bagged emasculated flowers) and natural selfing
were tested (N=10). Flowers were emasculated the day prior anthesis and
hand pollinated next day with a toothpick. Had-pollination was realized
with 3-4 gentle touches on the surface of the stigma. The mixes of
pollen were performed on an eppendorf based on the pollen counts maded
with Neubaeur chamber (each anther was counted 4 times for 20 different
anthers per species).

\textbf{Evolutive distance}

Two types of evolutive distances were calculated with MEGA7 thow kinds
of markers: 1) Internal transcribed spacer (ITS) and 2)
ribulose-bisphosphate carboxylase (RBCL)

\textbf{Traits}

Several traits of the ten species were measured. Pollen per anther was
counted, number of ovules, stigma width and length and stigmatic area,
style width and length, ovary width and length. Moreover stigma type was
tested. Self-incompatibility was

@@

@@

\hypertarget{ref-arceo2016}{}
Arceo-Gómez, G., and T.-L. Ashman. 2016. Invasion status and
phylogenetic relatedness predict cost of heterospecific pollen receipt:
Implications for native biodiversity decline. Journal of Ecology
104:1003--1008.
